---
title: "Network Analyses"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
    theme: flatly
    highlight: tango
    df_print: paged
---

```{r, echo = FALSE, message = FALSE, warning = FALSE }
knitr::opts_chunk$set(
  echo    = FALSE,    # hide code
  message = FALSE,    # hide messages
  warning = FALSE,    # hide warnings
  dpi     = 96,       # sharp figures
  fig.width = 6,      
  fig.height = 6,     
  dev     = "png"
)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE }

library(igraph)
library(ggraph)
library(modelsummary)

source("scripts_R/utils/check_pkgs.R")
source("scripts_R/utils/summarize_networks.R")
source("scripts_R/data_wrangling.R")
```

## 1. Network overview {.tabset}

Let's get an idea of the school network. Here I am using the data from the file "edgelist.csv", where I have directed nominations for 2 categories: "good friends" and "best friends". I build 2 networks: 1 for "good friends" (Amistad = 1), and one for "best friends" (Amistad = 2). Below, I plot the 2 networks and summarize the main properties in a table and in 3 histograms.

Node size reflects in-degree (the larger the node, the more nominations received)

### Good friends

```{r}

edge_list <- read_csv('data/edgelist.csv')  %>% 
  filter(status == "Amistad",
         weight == 1)

g1 <- graph_from_data_frame(d = edge_list, directed = TRUE)

good_friend <- summarize_network(g1) %>% 
  mutate(GoodFriend = Value) %>% 
  select(Metric, GoodFriend)
# --- Calculate In-degree ---
# Use mode = "in" to get the count of incoming edges for each node
node_in_degrees <- degree(g1, mode = "in")

# --- Scale Node Sizes (Recommended) ---
# Raw degrees might result in too small/large nodes. Scale them for better visuals.
# Example: base size 5 + degree * scaling factor 8
node_sizes <- (node_in_degrees * .4)
coords <- layout_with_fr(g1) # Calculate once

# --- Basic Plotting with Size based on In-degree ---
good_friend_plot <- plot(
  g1,
  layout = coords, # Use a suitable layout
  vertex.size = node_sizes,       # Set vertex size based on calculated in-degree
  edge.arrow.size = 0.1,
  vertex.label =NA,         # Might need smaller labels if nodes are large
  main = ""
)

good_friend_plot
```

### Best friends

```{r}

edge_list <- read_csv('data/edgelist.csv')  %>% 
  filter(status == "Amistad",
         weight == 2)

g2 <- graph_from_data_frame(d = edge_list, directed = TRUE)

# --- Calculate In-degree ---
# Use mode = "in" to get the count of incoming edges for each node
node_in_degrees <- degree(g2, mode = "in")

# --- Scale Node Sizes (Recommended) ---
# Raw degrees might result in too small/large nodes. Scale them for better visuals.
# Example: base size 5 + degree * scaling factor 8
node_sizes <- (node_in_degrees * .8)
coords <- layout_with_fr(g2) # Calculate once

best_friend_plot <- plot(
  g2,
  layout = coords, # Use a suitable layout
  vertex.size = node_sizes,       # Set vertex size based on calculated in-degree
  edge.arrow.size = 0.1,
  vertex.label =NA,         # Might need smaller labels if nodes are large
  main = ""
)

best_friend <- summarize_network(g2) %>% 
  mutate(BestFriend = Value) %>% 
  select(Metric, BestFriend)

best_friend_plot
```

## Table and visualization of network properties {.tabset}

Table summarizes network properties, plots show the 3 metrics that will later be used in the analyses: 1) In-degree, 2) Eigenvector Centrality, 3) Betweenness Centrality.

### Table

```{r}

nets <- merge(good_friend, best_friend, by = "Metric")
knitr::kable(nets, caption = "Network Summary")

```

### Metrics Network Good Friends

```{r, fig.width = 7, fig.height = 3 }

# 2. Compute centrality measures
in_deg    <- degree(g1, mode = "in")               # in-degree
eigen_c   <- eigen_centrality(g1, directed = TRUE)$vector
betw      <- betweenness(g1, directed = TRUE, normalized = TRUE)

# pack into a data.frame
df <- data.frame(
  in_degree  = in_deg,
  eigen_c    = eigen_c,
  betweenness= betw
)

# 3. Plot 1: in-degree distribution
p1 <- ggplot(df, aes(x = in_degree)) +
  geom_histogram(bins = 30, fill = "skyblue") +
    geom_vline(aes(xintercept = mean(in_degree, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +
  labs(
    title   = "In-degree",
    x       = "In-degree",
    y       = "Count"
  ) 

# 4. Plot 2: eigenvector centrality distribution
p2 <- ggplot(df, aes(x = eigen_c)) +
  geom_histogram(bins = 30, fill = "lightgreen") +
    geom_vline(aes(xintercept = mean(eigen_c, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +
  labs(
    title   = "Eigenvector Centrality",
    x       = "Eigenvector Centrality",
    y       = "Count"
  ) 

# 5. Plot 3: betweenness distribution
p3 <- ggplot(df, aes(x = betweenness)) +
  geom_histogram(bins = 30, fill = "salmon") +
      geom_vline(aes(xintercept = mean(betweenness, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +

  labs(
    title   = "Betweennesss Centrality",
    x       = "Betweennesss Centrality",
    y       = "Count"
  ) 

p1+p3+p2+
    plot_annotation(
    title = "Good friends network",
    theme = theme(plot.title = element_text(size = 14, face = "bold"))
  )

```

### Metrics Network Best Friends

```{r, fig.width = 7, fig.height = 3 }

# 2. Compute centrality measures
in_deg    <- degree(g2, mode = "in")               # in-degree
eigen_c   <- eigen_centrality(g2, directed = TRUE)$vector
betw      <- betweenness(g2, directed = TRUE, normalized = TRUE)

# pack into a data.frame
df <- data.frame(
  in_degree  = in_deg,
  eigen_c    = eigen_c,
  betweenness      <- betw
)

# 3. Plot 1: in-degree distribution
p1 <- ggplot(df, aes(x = in_degree)) +
  geom_histogram(bins = 30, fill = "skyblue") +
    geom_vline(aes(xintercept = mean(in_degree, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +
  labs(
    title   = "In-degree",
    x       = "In-degree",
    y       = "Count"
  ) 

# 4. Plot 2: eigenvector centrality distribution
p2 <- ggplot(df, aes(x = eigen_c)) +
  geom_histogram(bins = 30, fill = "lightgreen") +
      geom_vline(aes(xintercept = mean(eigen_c, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +
  labs(
    title   = "Eigenvector Centrality",
    x       = "Eigenvector Centrality",
    y       = "Count"
  ) 

# 5. Plot 3: betweenness distribution
p3 <- ggplot(df, aes(x = betweenness)) +
  geom_histogram(bins = 30, fill = "salmon") +
      geom_vline(aes(xintercept = mean(betweenness, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +
  labs(
    title   = "Betweennesss Centrality",
    x       = "Betweennesss Centrality",
    y       = "Count"
  ) 

p1+p3+p2+
    plot_annotation(
    title = "Best friends network",
    theme = theme(plot.title = element_text(size = 14, face = "bold"))
  )
```

## 2. Analyses on participants who are in the experiment {.tabset}

We now examine the network metrics for participants in the experiment. It may also be useful to compare different methods for computing these metrics.

Another thing to note is that the distribution of eigenvector centrality contains individuals with high values (EC or BC), which may affect the results.

In other words, differences in network metrics may not directly correspond to differences in appropriateness ratings as a function of these metrics.

### Good friends network

```{r, fig.width = 7, fig.height = 3 }
p1 <- data %>% 
  ggplot(aes(x = in_degree_1p)) +
  geom_bar(bins = 30, fill = "skyblue") +
  geom_vline(aes(xintercept = mean(in_degree_1p, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "In-degree") 

p2 <- data %>% 
  ggplot(aes(x = `BC+I`)) +
  geom_histogram(bins = 30, fill = "salmon") +
    geom_vline(aes(xintercept = mean(`BC+I`, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "Betweenness Centrality") 

p3 <- data %>% 
  ggplot(aes(x = `EC+I(in)`)) +
  geom_histogram(bins = 30, fill = "lightgreen") +
    geom_vline(aes(xintercept = mean(`EC+I(in)`, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1)+
  labs(title = "Eigenvector Centrality") 

(p1+p2+p3) +
    plot_annotation(
    title = "Good friends network",
    theme = theme(plot.title = element_text(size = 14, face = "bold"))
  )
```

### Best Friends Network

```{r, fig.width = 7, fig.height = 3 }

p4 <- data %>% 
  ggplot(aes(x = in_degree_2p)) +
  geom_bar(bins = 30, fill = "skyblue") +
    geom_vline(aes(xintercept = mean(in_degree_2p, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "In-degree") 

p5 <- data %>% 
  ggplot(aes(x = `BC+II`)) +
  geom_histogram(bins = 30, fill = "salmon") +
    geom_vline(aes(xintercept = mean(`BC+II`, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "Betweenness Centrality") 

p6 <- data %>% 
  ggplot(aes(x = `EC+II(in)`)) +
  geom_histogram(bins = 30, fill = "lightgreen") +
    geom_vline(aes(xintercept = mean(`EC+II(in)`, na.rm = TRUE)), color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "Eigenvector Centrality") 

(p4+p5+p6) +
      plot_annotation(
    title = "Best friends network",
    theme = theme(plot.title = element_text(size = 14, face = "bold"))
   # subtitle = "Rating: 0 = totally inappropriate, 5 - totally appropriate"
  )
```

## 3. Regression analyses of appropriateness of violation/punishment of violation

Using the above mentioned network metrics as independent variables, I will not explore weather they predict any difference in the appropriate ratings for both **violations** and **punishment of violations**, in the two treatments (friends vs. non-friends).

I will start with a simple regression of rating \~ [network metric].

I will use both metrics from the "good-friends" and the "best-friends" social networks. In-degree metrics were standardized for easier interpretation.

## 3.1 Appropriateness ratings by network measures {.tabset}

The only significant predictors of ratings are **in-degree** and **eigenvector** centrality, both with a positive effect. This shows that [participants who are more central rate violations as more appropriate.]{.underline}

### In-degree [Good-friends]

```{r}
tab_model(lmer(rating ~ std_in_degree_1p   + (1|ID2), data = net_appropriateness))
```

### In-degree [Best-friends]

```{r}
tab_model(lmer(rating ~ std_in_degree_2p   + (1|ID2), data = net_appropriateness))
```

### Betweenness Centrality [Good-friends]

```{r}
tab_model(lmer(rating ~  BC1  + (1|ID2), data = net_appropriateness))
```

### Betweenness Centrality [Best-friends]

```{r}
tab_model(lmer(rating ~  BC2  + (1|ID2), data = net_appropriateness))
```

### Eigenvector Centrality [Good-friends]

```{r}
tab_model(lmer(rating ~  EC1  + (1|ID2), data = net_appropriateness))
```

### Eigenvector Centrality [Best-friends]

```{r}
tab_model(lmer(rating ~  EC2  + (1|ID2), data = net_appropriateness))
```

## 3.2 Appropriateness ratings by network measures and treatment {.tabset}

I know include treatment (friend vs. non-friend) as an interaction in the regression model. We again see a main effect of ... and a main effect of treatment, as we saw in the analyses without network propertis.

This means that [ratings of approppriateness of a violation of a friend vs. a non-friend were not related with the centrality of the rater.]{.underline}

### In-degree [Good-friends]

```{r}
## violation ~ net * treatment
tab_model(lmer(rating ~ in_degree_1p * treatment  + (1|ID2), data = net_appropriateness))
```

### In-degree [Best-friends]

```{r}
tab_model(lmer(rating ~ std_in_degree_2p * treatment  + (1|ID2), data = net_appropriateness))
```

### Betweenness Centrality [Good-friends]

```{r}
tab_model(lmer(rating ~  BC1 * treatment  + (1|ID2), data = net_appropriateness))
```

### Betweenness Centrality [Best-friends]

```{r}
tab_model(lmer(rating ~  BC2  * treatment + (1|ID2), data = net_appropriateness))
```

### Eigenvector Centrality [Good-friends]

```{r}
tab_model(lmer(rating ~  EC1  * treatment + (1|ID2), data = net_appropriateness))
```

### Eigenvector Centrality [Best-friends]

```{r}
tab_model(lmer(rating ~  EC2  * treatment + (1|ID2), data = net_appropriateness))
```

```{r}

```

## 4 Punishment {.tabset}

We now turn to the appropriateness of punishment. Remember here that the rater is not the one punishing, but rather they are rating how appropriate it is for a friend, to punish a friend/non-friends.

## 4.1 Appropriatness of punishment by network measures {.tabset}

First we only look at appropriateness of punishment as a function of network metrics.

```{r}
net_punishment <- merge(network_info, data_punish_all, by = "ID2")

```

### In-degree [Good-friends]

```{r}
tab_model(lmer(rating ~ std_in_degree_1p  + (1|ID2), data = net_punishment))
```

### In-degree [Best-friends]

```{r}
tab_model(lmer(rating ~ std_in_degree_2p  + (1|ID2), data = net_punishment))
```

### Betweennesss Centrality [Good-Friends]

```{r}
tab_model(lmer(rating ~ BC1 + (1|ID2), data = net_punishment))
```

### Betweennesss Centrality [Best-Friends]

```{r}
tab_model(lmer(rating ~ BC2 + (1|ID2), data = net_punishment))
```

### Eigenvector Centrality [Good-Friends]

```{r}
tab_model(lmer(rating ~ EC1 + (1|ID2), data = net_punishment))
```

### Eigenvector Centrality [Best-Friends]

```{r}
tab_model(lmer(rating ~ EC2 + (1|ID2), data = net_punishment))
```

## 4.2 Appropriatness of punishment by network measures \* treatment {.tabset}

### In-degree [Good-Friends]

```{r}
tab_model(lmer(rating ~ std_in_degree_1p * treatment  + (1|ID2), data = net_punishment))
```

### In-degree [Best-Friends]

```{r}
tab_model(lmer(rating ~ std_in_degree_2p * treatment + (1|ID2), data = net_punishment))
```

### Betweenness Centrality [Good-Friends]

```{r}
tab_model(lmer(rating ~ BC1 * treatment  + (1|ID2), data = net_punishment))
```

### Betweenness Centrality [Best-Friends]

```{r}
tab_model(lmer(rating ~ BC2 * treatment  + (1|ID2), data = net_punishment))
```

### Eigenvector Centrality [Good-Friends]

```{r}
tab_model(lmer(rating ~ EC1 * treatment   + (1|ID2), data = net_punishment))
```

### Eigenvector Centrality [Best-Friends]

```{r}
tab_model(lmer(rating ~ EC2 * treatment  + (1|ID2), data = net_punishment))
```

## 3-level interaction

```{r}
tab_model(lmer(rating ~ EC1 * treatment *domain  + (1|ID2), data = net_punishment))

```

## Vizualization of main regression results {.tabset}

Showing the most interesting results across all regressions

### Violation rating \~ In-degree

```{r}
net_appropriateness %>% 
  group_by(ID2, treatment) %>% 
  mutate(mean_rating = mean(rating),
         EC1 = EC1) %>% 
  select(mean_rating, ID2, in_degree_1p,domain) %>% 
  distinct() %>% 
  ggplot() +
  geom_point(aes(y=mean_rating, x= in_degree_1p, color=treatment), alpha = .2)+
  geom_smooth(aes(y=mean_rating, x= in_degree_1p, color=treatment), method = 'lm')+
  scale_fill_brewer(palette = "Set2")  + 
  scale_color_brewer(palette = "Set2") +
  labs(title="Appropriateness of violations by in-degree")
```

The higher participants' in degree, the higher the appropriateness rating. Ratings are averaged across the 5 domain for each participants: Each dot is 1 participant in 1 treatment (nf = non-friend, f = friend)

### Violation rating \~ Eigenvector Centrality

```{r}
net_appropriateness %>% 
  group_by(ID2, treatment) %>% 
  mutate(mean_rating = mean(rating),
         EC1 = EC1) %>% 
  select(mean_rating, ID2, in_degree_1p, EC1) %>% 
  distinct() %>% 
  ggplot() +
  geom_point(aes(y=mean_rating, x= EC1, color=treatment), alpha = .2)+
  geom_smooth(aes(y=mean_rating, x= EC1, color=treatment), method = 'lm')+
  scale_fill_brewer(palette = "Set2")  + 
  scale_color_brewer(palette = "Set2") +
  labs(title="Appropriateness of violations by Eigenvector Centrality")
```

The higher participants' in degree, the higher the appropriateness rating. Ratings are averaged across the 5 domain for each participants: Each dot is 1 participant in 1 treatment (nf = non-friend, f = friend). Note that most participants are between 0 and 0.02

### Punishment Rating \~ Eigenvector Centrality

```{r}
net_punishment %>% 
  group_by(ID2, treatment) %>% 
  mutate(mean_rating = mean(rating),
         EC1 = EC1) %>% 
  select(mean_rating, ID2, EC1, domain) %>% 
  distinct() %>% 
 ggplot() +
  geom_point(aes(y=mean_rating, x= EC1, color = treatment), alpha = .2) +
  geom_smooth(aes(y=mean_rating, x= EC1, color=treatment), method = 'lm')+
  scale_fill_brewer(palette = "Set2")  + 
  scale_color_brewer(palette = "Set2")  
```

### Punishment Rating \~ Eigenvector Centrality \* Treatment \* Domain

```{r}
net_punishment %>% 
  group_by(ID2, treatment, domain) %>% 
  mutate(mean_rating = mean(rating),
         EC1 = EC1) %>% 
  select(mean_rating, ID2, EC1, domain) %>% 
  distinct() %>% 
 ggplot() +
  geom_point(aes(y=mean_rating, x= EC1, color = treatment), alpha = .4) +
  geom_smooth(aes(y=mean_rating, x= EC1, color=treatment), method = 'lm')+
  scale_fill_brewer(palette = "Set2")  + 
  scale_color_brewer(palette = "Set2")+
facet_wrap(~domain)
```
