---
title: "Scenarios Analyses"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
    theme: flatly
    highlight: tango
    df_print: paged
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(
  echo    = FALSE,    # hide code
  message = FALSE,    # hide messages
  warning = FALSE,    # hide warnings
  fig.width = 6,      
  fig.height = 6,     
  dev     = "png"
)

```

```{r, include=FALSE}

## change options to only run linear contrast for factors
options(contrasts = c("contr.treatment", "contr.treatment"))

## prepare environment
source("scripts_R/utils/check_pkgs.R")
source("scripts_R/utils/make_plots.R")
source("scripts_R/data_wrangling.R")

```

```{r}

```

## 1. Overview of scenario data {.tabset}

First, I look at how participants rated the appropriateness of each violation in each scenario, and the appropriateness of the different types of punishments in that scenario.

[Data is displayed separately by treatment]{.underline}, i.e. whether the violation is done by a **friend/non-friend**, and weather the punishment is performed by a friend to a **friend/non-friend**. Treatments are between subjects.

**Insights:**\
All violations are considered fairly inappropriate, with insulting a family member and stealing being the worst.\
There is quite some variation in appropriateness of the different types of punishment.\
There are no striking differences between treatments.

### Stealing

```{r,  fig.width = 6, fig.height = 6  }

## appropriateness of behavious
stealapp <- faceted_barplot(
  data = data,
  x_var = stealapp,
  facet_var = stealfriend,
  title_text = "Appropriateness") +
  theme(plot.background = element_rect(fill = "grey90"))+
  facet_wrap( ~ stealfriend, labeller = labeller(.default = facet_labels)) 
  
## reactions

#1.doing nothing
stealnoth <- faceted_barplot(
  data = data,
  x_var = stealnoth,
  facet_var = stealfriend,
  title_text = "Punishment: do nothing"
) +
  facet_wrap( ~ stealfriend, labeller = labeller(.default = facet_labels))

#2.angry remark

stealremark <- faceted_barplot(
  data = data,
  x_var = stealremark,
  facet_var = stealfriend,
  title_text = "Punishment: angry remark"
) +
 labs() +
  facet_wrap( ~ stealfriend, labeller = labeller(.default = facet_labels))

#3.gossip
stealtalk <- faceted_barplot(
  data = data,
  x_var = stealtalk,
  facet_var = stealfriend,
  title_text = "Punishment: gossip"
) + facet_wrap( ~ stealfriend, labeller = labeller(.default = facet_labels))

#4.avoid
stealavoid <- faceted_barplot(
  data = data,
  x_var = stealavoid,
  facet_var = stealfriend,
  title_text = "Punishment: avoid"
) +

  facet_wrap( ~ stealfriend, labeller = labeller(.default = facet_labels))

#5. physical punishment
# faceted_barplot(
#   data = data,
#   x_var = stealphyspunapp,
#   facet_var = stealfriend,
#   title_text = "Stealing: physical punishment"
# ) + facet_wrap( ~ stealfriend, labeller = labeller(.default = facet_labels))


panel_punishment <- stealapp / (stealnoth + stealavoid)  / (stealtalk   + stealremark) +
  plot_annotation(
    title = "Stealing",
    theme = theme(plot.title = element_text(size = 14, face = "bold")),
    subtitle = "Rating: 0 = totally inappropriate, 5 - totally appropriate"
  )

panel_punishment
    
```

### Phone

```{r}

## appropriateness of behavious
phoneapp <- faceted_barplot(
  data = data,
  x_var = phoneapp,
  facet_var = phonefriend,
  title_text = "Appropriateness"
) +  facet_wrap( ~ phonefriend, labeller = labeller(.default = facet_labels))+
    theme(plot.background = element_rect(fill = "grey90"))


## reactions
#1. doing nothing

phonenoth <- faceted_barplot(
  data = data,
  x_var = phonenoth,
  facet_var = phonefriend,
  title_text = "Punishment: do nothing"
) +
 # labs(subtitle = "Friend A does nothing about the person who behaved like the PURPLE triangle") +
  facet_wrap( ~ phonefriend, labeller = labeller(.default = facet_labels))

#2. angry remark
phoneremark <- faceted_barplot(
  data = data,
  x_var = phoneremark,
  facet_var = phonefriend,
  title_text = "Punishment: angry remark"
) +
 # labs(subtitle = "Friend B makes an angry remark to the person who behaved like the PURPLE triangle.") +
  facet_wrap( ~ phonefriend, labeller = labeller(.default = facet_labels))

#3. gossip
phonetalk <- faceted_barplot(
  data = data,
  x_var = phonetalk,
  facet_var = phonefriend,
  title_text = "Punishment: gossip"
) + 
#  labs(subtitle = "Friend C talks to someone else about the person who behaved like the PURPLE triangle. ") +
  facet_wrap( ~ phonefriend, labeller = labeller(.default = facet_labels))

#4.avoid
phoneavoid <- faceted_barplot(
  data = data,
  x_var = phoneavoid,
  facet_var = phonefriend,
  title_text = "Punishment: avoid"
) +
  #labs(subtitle = "Friend D makes a point of avoiding the person who behaved like the PURPLE triangle in the future, even when that person is not behaving in this way.") +
  facet_wrap( ~ phonefriend, labeller = labeller(.default = facet_labels))



panel_phone <- phoneapp / (phonenoth + phoneavoid)  / (phonetalk   + phoneremark) +
  plot_annotation(
    title = "Phone in class",
    theme = theme(plot.title = element_text(size = 14, face = "bold")),
    subtitle = "Rating: 0 = totally inappropriate, 5 - totally appropriate"
  )

panel_phone
    
```

### Interrupt

```{r}

## appropriateness of behavious
interruptapp <- faceted_barplot(
  data = data,
  x_var = interruptapp,
  facet_var = factor(interruptfriend),
  title_text = "Appropriateness"
) +  facet_wrap( ~ interruptfriend, labeller = labeller(.default = facet_labels)) +
  theme(plot.background = element_rect(fill = "grey90"))

## reactions
#1. doing nothing

interruptnoth <- faceted_barplot(
  data = data,
  x_var = interruptnoth,
  facet_var = interruptfriend,
  title_text = "Punishment: do nothing"
) +
 # labs(subtitle = "Friend A does nothing about the person who behaved like the PURPLE triangle") +
  facet_wrap( ~ interruptfriend, labeller = labeller(.default = facet_labels))

#2. angry remark

interruptremark <- faceted_barplot(
  data = data,
  x_var = interruptremark,
  facet_var = interruptfriend,
  title_text = "Punishment: angry remark"
) +
 # labs(subtitle = "Friend B makes an angry remark to the person who behaved like the PURPLE triangle.") +
  facet_wrap( ~ interruptfriend, labeller = labeller(.default = facet_labels))

#3. gossip

interrupttalk <- faceted_barplot(
  data = data,
  x_var = interrupttalk,
  facet_var = interruptfriend,
  title_text = "Punishment: gossip"
) + 
#  labs(subtitle = "Friend C talks to someone else about the person who behaved like the PURPLE triangle. ") +
  facet_wrap( ~ interruptfriend, labeller = labeller(.default = facet_labels))

#4.avoid

interruptavoid <- faceted_barplot(
  data = data,
  x_var = interruptavoid,
  facet_var = interruptfriend,
  title_text = "Punishment: avoid"
) +
  #labs(subtitle = "Friend D makes a point of avoiding the person who behaved like the PURPLE triangle in the future, even when that person is not behaving in this way.") +
  facet_wrap( ~ interruptfriend, labeller = labeller(.default = facet_labels))


panel_interrupt <- interruptapp / (interruptnoth + interruptavoid)  / (interrupttalk   + interruptremark) +
  plot_annotation(
    title = "Interrupt in class",
    theme = theme(plot.title = element_text(size = 14, face = "bold")),
    subtitle = "Rating: 0 = totally inappropriate, 5 - totally appropriate"
  )

panel_interrupt

```

### Insult

```{r}

## appropriateness of behaviors
insultapp <- faceted_barplot(
  data = data,
  x_var = insultapp,
  facet_var = insultfriend,
  title_text = "Appropriateness"
) +  facet_wrap( ~ insultfriend, labeller = labeller(.default = facet_labels)) +
   theme(plot.background = element_rect(fill = "grey90"))

## reactions

#1. doing nothing

insultnoth <- faceted_barplot(
  data = data,
  x_var = insultnoth,
  facet_var = insultfriend,
  title_text = "Punishment: do nothing"
) +
 # labs(subtitle = "Friend A does nothing about the person who behaved like the PURPLE triangle") +
  facet_wrap( ~ insultfriend, labeller = labeller(.default = facet_labels))

#2. angry remark

insultremark <- faceted_barplot(
  data = data,
  x_var = insultremark,
  facet_var = insultfriend,
  title_text = "Punishment: angry remark"
) +
 # labs(subtitle = "Friend B makes an angry remark to the person who behaved like the PURPLE triangle.") +
  facet_wrap( ~ insultfriend, labeller = labeller(.default = facet_labels))

#3. gossip

insulttalk <- faceted_barplot(
  data = data,
  x_var = insulttalk,
  facet_var = insultfriend,
  title_text = "Punishment: gossip"
) + 
#  labs(subtitle = "Friend C talks to someone else about the person who behaved like the PURPLE triangle. ") +
  facet_wrap( ~ insultfriend, labeller = labeller(.default = facet_labels))

#4.avoid

insultavoid <- faceted_barplot(
  data = data,
  x_var = insultavoid,
  facet_var = insultfriend,
  title_text = "Punishment: avoid"
) +
  #labs(subtitle = "Friend D makes a point of avoiding the person who behaved like the PURPLE triangle in the future, even when that person is not behaving in this way.") +
  facet_wrap( ~ insultfriend, labeller = labeller(.default = facet_labels))


panel_insult <- insultapp / (insultnoth + insultavoid)  / (insulttalk   + insultremark) +
  plot_annotation(
    title = "Insult family member",
    theme = theme(plot.title = element_text(size = 14, face = "bold")),
    subtitle = "Rating: 0 = totally inappropriate, 5 - totally appropriate"
  )

panel_insult

```

### Exclude

```{r}

## appropriateness of behaviors
onlineapp <- faceted_barplot(
  data = data,
  x_var = onlineapp,
  facet_var = onlinefriend,
  title_text = "Appropriateness"
) +  facet_wrap( ~ onlinefriend, labeller = labeller(.default = facet_labels)) +
   theme(plot.background = element_rect(fill = "grey90"))


#1. doing nothing

onlinenothing <- faceted_barplot(
  data = data,
  x_var = onlinenothing,
  facet_var = onlinefriend,
  title_text = "Punishment: do nothing"
) +
 # labs(subtitle = "Friend A does nothing about the person who behaved like the PURPLE triangle") +
  facet_wrap( ~ onlinefriend, labeller = labeller(.default = facet_labels))

#2. angry remark

onlineremark <- faceted_barplot(
  data = data,
  x_var = onlineremark,
  facet_var = onlinefriend,
  title_text = "Punishment: angry remark"
) +
 # labs(subtitle = "Friend B makes an angry remark to the person who behaved like the PURPLE triangle.") +
  facet_wrap( ~ onlinefriend, labeller = labeller(.default = facet_labels))

#3. gossip

onlinetalk <- faceted_barplot(
  data = data,
  x_var = onlinetalk,
  facet_var = onlinefriend,
  title_text = "Punishment: gossip"
) + 
#  labs(subtitle = "Friend C talks to someone else about the person who behaved like the PURPLE triangle. ") +
  facet_wrap( ~ onlinefriend, labeller = labeller(.default = facet_labels))

#4.avoid
onlineavoid <- faceted_barplot(
  data = data,
  x_var = onlineavoid,
  facet_var = onlinefriend,
  title_text = "Punishment: avoid"
) +
  #labs(subtitle = "Friend D makes a point of avoiding the person who behaved like the PURPLE triangle in the future, even when that person is not behaving in this way.") +
  facet_wrap( ~ onlinefriend, labeller = labeller(.default = facet_labels))

panel_online <- onlineapp / (onlinenothing + onlineavoid)  / (onlinetalk   + onlineremark) +
  plot_annotation(
    title = "Exclude from chat",
    theme = theme(plot.title = element_text(size = 14, face = "bold")),
    subtitle = "Rating: 0 = totally inappropriate, 5 - totally appropriate"
  )

panel_online

```

## 2. Analysis of appropriateness of violations across all domains {.tabset}

Now I merge all the scenarios together.

**Insight:**

Overall appropriateness is quite low, with again no striking difference between treatments.

```{r, fig.width = 6, fig.height = 4}
data_app_all %>% 
  ggplot() +
  geom_bar(aes(x = rating_f, fill = domain)) +
  scale_x_discrete(drop = FALSE) +  # show all factor levels
  facet_wrap( ~ treatment) +
  labs(title = "Appropriateness of behaviours across domains",
       x = "rating") +
  theme(strip.background = element_rect(fill = "lightgrey"))
```

## 2.1 ANOVA

First I run an ANOVA to get an idea of differences across treatments and domains. The ANOVA showed that both treatment and domain significantly affected ratings, with a significant interaction indicating that the effect of treatment varied across domains.

```{r}
summary(aov(lm(rating ~ treatment * domain , data = data_app_all)))
```

## 2.2 Regressions {.tabset}

I now run two multilevel linear regressions with participants as random intercept to zoom in on size and direction of efffects.

First, a linear regression predicting rating \~ treatment shows a small effect of treatment: [violations by friends are considered slightly more appropriate.]{.underline}

If we include a test of interactions between treatment and domain, we can see that the effect is driven by the "exclude from online chat scenario:

>  "One student who is part of your **friends/others** **group** (**Student A**) excluded **another student**, who is also a part of your **friends group**, from a group chat on WhatsApp."

### Rating \~ Treatment

```{r}
sjPlot::tab_model(lme4::lmer(rating ~ treatment + (1|ID2), data = data_app_all))
```

### Rating \~ Treatment \* Domain

```{r}

sjPlot::tab_model(lme4::lmer(rating ~ treatment * domain + (1|ID2), data = data_app_all))


# library(brms)
# 
# fancymodel <- brm(
#   formula = rating_f ~ treatment,
#   data = data_app_all,
#   family = cumulative(link = "logit"),
#   chains = 4, cores = 4, iter = 4000
# )
# 
# fancymodel_interaction <- brm(
#   formula = rating_f ~ treatment * domain,
#   data = data_app_all,
#   family = cumulative(link = "logit"),
#   chains = 4, cores = 4, iter = 4000
# )
# 
# sjPlot::plot_model(fancymodel_interaction)
# 
# pp_check(fancymodel_interaction)

```

## 3. Analysis of appropriateness of punishment if friend vs. non-friend

Now I look at the appropriateness ratings of punishments if the violator is a friend vs. non friend.

## 3.1 Vizualization {.tabset}

Again, visual inspection does not reveal any striking differences bewteen treatments and domains.

### Overall appropriatess across domains

```{r, fig.width = 6, fig.height = 4 }

data_punish_all %>% 
  ggplot() +
  geom_bar(aes(x = rating_f, fill = domain)) +
  scale_x_discrete(drop = FALSE) +  # show all factor levels
  facet_wrap( ~ treatment) +
  labs(title = "Appropriateness of punishment across domains") +
  theme(strip.background = element_rect(fill = "lightgrey"))
```

### Appropriateness by domain

```{r, fig.width = 6, fig.height = 4 }
data_punish_all %>% 
  ggplot(aes(x = treatment, y = rating, fill = treatment)) +
  geom_half_boxplot() +
  geom_half_point(aes(color = treatment), alpha = .2) +
  labs(title = "Appropriateness of punishment in each domain") +
  facet_wrap(~domain) +
  theme(strip.background = element_rect(fill = "lightgrey")) +
  scale_fill_brewer(palette = "Set2")  + 
  scale_color_brewer(palette = "Set2")

```

### Appropriateness of punishment by treatment and punisment type

```{r, fig.width = 6, fig.height = 4 }

data_punish_all %>% 
  ggplot(aes(x = treatment, y = rating, fill = punishment_type)) +
  geom_half_boxplot() +
  geom_half_point(aes(color = punishment_type), alpha = .2) +
  labs(title = "Appropriateness of punishment in each domain") +
 # facet_wrap(~domain) +
  theme(strip.background = element_rect(fill = "lightgrey")) +
  scale_fill_brewer(palette = "Set2")  + 
  scale_color_brewer(palette = "Set2")
```

## 3.1 ANOVA

First, I run an ANOVA to get an idea of differences in mean ratings across treatments and domains.

There was **no significant main effect of treatment**.

In contrast, **domain** was significant, suggesting that the ratings varied meaningfully across different domains.

The interaction between **treatment and domain** was also significant, implying that the effect of treatment differed depending on the domain.

```{r}

summary(aov(lm(rating ~ treatment * domain , data = data_punish_all)))

```

## 3.2 Regressions {.tabset}

Now I fit 3 linear regressions to better explore directions of effects.

First, only with treatment as independent variable, then I add interactions with domain (insult as reference) and finally I also include punishment type as exploratory variable (angry_remark as reference).

The main results are [see tabs below for full results]:

**Model 1: no effect**

**Model2 :**

-   Raters considered [more appropriate to punish friends who exclude others from group chats.]{.underline}

**Model 3:**

-   The domain "texting" was associated with significantly lower ratings compared to the reference domain.
-   The punishment types "do nothing" and "gossip" led to significantly lower ratings compared to the reference punishment type.

There a few significant interactions:

-   The combination of "interrupt" domain and "avoid" punishment type was associated with lower ratings compared to the reference groups.

-   The combination of "online" domain and "do nothing" punishment type was associated with lower ratings compared to the reference groups.

-   The combination of "stealing" domain and "do nothing" punishment type was associated with lower ratings compared to the reference groups.

-   The combination of "texting" domain and "do nothing" punishment type was associated with higher ratings compared to the reference groups.

-    The combination of "stealing" domain and "gossip" punishment type was also associated with higher ratings compared to the reference groups.

### Rating \~ Treatment

```{r, fig.width = 6, fig.height = 4 }
sjPlot::tab_model(lme4::lmer(rating ~ treatment  + (1|ID2), data = data_punish_all))
```

### Rating \~ Treatment \* Domain

```{r, fig.width = 4, fig.height = 6 }
sjPlot::tab_model(lme4::lmer(rating ~ treatment * domain + (1|ID2), data = data_punish_all))
```

### Rating \~ Treatment \* Domain\* punishment type

```{r, fig.width = 4, fig.height = 6 }
sjPlot::tab_model(lme4::lmer(rating ~ treatment * domain *punishment_type + (1|ID2), data = data_punish_all))
```
